# Module 1: Introduction to the Course
***


**Module Sections:**

* Welcome to the Course
* Introduction to R
* Introductory Lecture - Data is Beautiful, Insightful, Powerful, Deceitful
* Finger Exercises 
* Module 1: Homework

Module Content:

* [Module 1 Slides](./files/M1/Lecture_Slides_01.pdf)
* [Homework 1 Background Paper -  The Persistent Effects Of Peru’s Mining Mita](./files/M1/M1Paper.pdf) 
* [R Instructions](./files/M1/R_Instructions.pdf)  
* [Intro to R Zip File](./files/M1/14_310x_Intro_to_R_.zip.pdf)  
* [Citations Data for Homework 1]("./files/CitesforSara.csv")


## Introduction to R

First we setup the environment and install the course files

```{r, eval = FALSE}
library(swirl)
install_course_zip("./files/M1/14_310x_Intro_to_R_.zip",multi=FALSE)
swirl()
```

### Module 1 Homework

This is a sample of some of the homework answers. Some questions were observational or required interpretation of maps for example, as such these are not inluded here.

```{r}
library(tidyverse)
papers <- as_tibble(read_csv("./files/M1/CitesforSara.csv"))

```

Q. 19 Let's take a look at some of the most popular papers. 
Using the filter() method, how many records exist where there are greater than or equal to 100 citations?

```{r}

#First lets look at our data
head(papers)
arrange(papers,desc(cites), title)

papers %>%
  filter(cites >= 100) 
```

Q.20 Use the group_by() function to group papers by journal. How many total citations exist for the journal "Econometrica"?

```{r}
papers %>%
  group_by(journal) %>%
  summarise(sum(cites))

#or

summarize(group_by
          (papers, journal), 
          SumOfCites = sum(cites))
```

Q.21 How many distinct primary authors (au1) exist in this dataset?
  
```{r}
papers %>%
  summarise(n_distinct(au1))

#or

n_distinct(papers$au1)
```
  
